Analysis-ready VCF at Biobank scale using Zarr

### 2024-06-07 paper.tex
Apply Jeff's comments to intro

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Please note that whilst this template provides a
% preview of the typeset manuscript for submission, it
% will not necessarily be the final publication layout.
%
% letterpaper/a4paper: US/UK paper size toggle
% num-refs/alpha-refs: numeric/author-year citation and bibliography toggle

@@

To achieve this full potential we 
need a new generation of tools that operate directly
on a primary data representation that supports 
efficient access across a range of applications,
with native support for cloud object storage.
Such a representation can be termed ``analysis-ready''
and ``cloud-native''~\citep{abernathey2021cloud}.
For the representation to be FAIR~\citep{wilkinson2016fair},
it must also be \emph{accessible}, using protocols that are 
``open, free, and universally implementable''.
There is currently no efficient, FAIR representation of genetic variation
data suitable for cloud deployments.
Hail~\cite{ganna2016ultra,hail2024} has become the dominant platform
for quality control of large-scale variation datasets, 
and has been instrumental in projects such as 
gnomadAD~\cite{karczewski2020mutational,chen2024genomic}.
Glow~\cite{glow2024} has recently emerged as a similar approach
to performing distributed analytics on genetic variation data.
While Hail and Glow are built on open components
from the Hadoop distributed computing ecosystem~\citep{white2012hadoop},
they are both complex projects consisting of 
tightly coupled components that are not intended for external reuse.
Similarly, commercial solutions that have emerged to facilitate
the analysis of large-scale genetic variation data are either
based on proprietary~\cite{basespace2024,graf2024,googlelifesciences2024,
awshealthomics2024,microsoftgenomics2024}
or single-vendor technologies~\cite[e.g.][]{tiledb2024,genomicsdb2024}.
The next generation of VCF analysis methods requires
an open, free and transparent data representation 
with multiple independent implementations.

>>>>

%,%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Please note that whilst this template provides a
% preview of the typeset manuscript for submission, it
% will not necessarily be the final publication layout.
%
% letterpaper/a4paper: US/UK paper size toggle
% num-refs/alpha-refs: numeric/author-year citation and bibliography toggle

@@

To achieve this full potential we 
need a new generation of tools that operate directly
on a primary data representation that supports 
efficient access across a range of applications,
with native support for cloud object storage.
Such a representation can be termed ``analysis-ready''
and ``cloud-native''~\citep{abernathey2021cloud}.
For the representation to be FAIR~\citep{wilkinson2016fair},
it must also be \emph{accessible}, using protocols that are 
``open, free, and universally implementable''.
There is currently no efficient, FAIR representation of genetic variation
data suitable for cloud deployments.
Hail~\cite{ganna2016ultra,hail2024} has become the dominant platform
for quality control of large-scale variation datasets, 
and has been instrumental in projects such as 
gnomadAD~\cite{karczewski2020mutational,chen2024genomic}.
% Glow~\cite{glow2024} has recently emerged as a similar approach
% to performing distributed analytics on genetic variation data.
While Hail is built on open components
from the Hadoop distributed computing ecosystem~\citep{white2012hadoop},
the details of its MatrixTable format are not documented
or intended for external reuse. 
% cite? https://dev.hail.is/t/matrixtable-file-format-reference/173/6
Similarly, commercial solutions that have emerged to facilitate
the analysis of large-scale genetic variation data are either
based on proprietary~\cite{basespace2024,graf2024,googlelifesciences2024,
awshealthomics2024,microsoftgenomics2024}
or single-vendor technologies~\cite[e.g.][]{tiledb2024,genomicsdb2024}.
The next generation of VCF analysis methods requires
an open, free and transparent data representation 
with multiple independent implementations.
